---
title: "hw05CarolineSeip"
author: "Caroline Seip"
date: "October 16, 2018"
output: github_document
---
#Table of Contents

#Load packages

Load the required packages for this assignment:

```{r}
suppressPackageStartupMessages(library(tidyverse))      
suppressPackageStartupMessages(library(knitr)) 
suppressPackageStartupMessages(library(gapminder)) 
suppressPackageStartupMessages(library(plotly)) 
```

#Part 1: Factor management

In this section we will determine whether we are working with factors, then drop factors and reorder the levels.

First let's check if we are working with factors:

```{r}
glimpse(gapminder)
```

Ok, so now we know that country and continent are both factors, and there are 1704 observations in the gapminder dataset.

Let's drop the continent Oceania from the gapminder dataset now:

##Drop Oceania

```{r}
noO <- gapminder %>% 
  filter(continent != "Oceania") %>% #filter gapminder dataset for continents that are not Oceania
  glimpse()
```

Now we can see that the tibble only contains 1680 observations. So it worked! We removed Oceania.

Let's check the levels of the continent factor of our new object:

```{r}
unique(noO$continent) #check levels of continent
```

Uh oh! There are no values for Oceania in the new object, but it still exists as a level. Let's get rid of it.

##Remove unused factor levels

Using `fct_drop` let's remove the unused factor level:

```{r}
dropO <- noO %>% #make a new variable
  mutate(continent = fct_drop(continent)) #drop unused factors of continent

unique(dropO$continent) #check levels of continent in the new variable
```

Now we can see that Oceania was removed as a factor, yay!

#Reorder the levels of country

For this part I will reorder the levels of continent by minimum life expectancy using `fct_reorder`.

```{r}
gapminder %>% 
  #Make a plot, reordering continent by minimum life expectancy
  ggplot(aes(fct_reorder(continent, lifeExp, .fun = min), lifeExp, fill = continent)) +
  #Add a violin plot
  geom_violin() +
  #Add an x label
  xlab("Continent") +
  #Add a y label
  ylab("Life expectancy (years)")
```

Yay! We reordered the continents by minimum life expectancy.



