---
title: "hw05CarolineSeip"
author: "Caroline Seip"
date: "October 16, 2018"
output: github_document
always_allow_html: yes
---
#Table of Contents

#Load packages

Load the required packages for this assignment:

```{r}
suppressPackageStartupMessages(library(tidyverse))      
suppressPackageStartupMessages(library(knitr)) 
suppressPackageStartupMessages(library(gapminder)) 
suppressPackageStartupMessages(library(plotly)) 
suppressPackageStartupMessages(library(htmlwidgets))
```

#Part 1: Factor management

In this section we will determine whether we are working with factors, then drop factors and reorder the levels.

First let's check if we are working with factors:

```{r}
glimpse(gapminder)
```

Ok, so now we know that country and continent are both factors, and there are 1704 observations in the gapminder dataset.

Let's drop the continent Oceania from the gapminder dataset now:

##Drop Oceania

```{r}
noO <- gapminder %>% 
  filter(continent != "Oceania") %>% #filter gapminder dataset for continents that are not Oceania
  glimpse()
```

Now we can see that the tibble only contains 1680 observations. So it worked! We removed Oceania.

Let's check the levels of the continent factor of our new object:

```{r}
unique(noO$continent) #check levels of continent
```

Uh oh! There are no values for Oceania in the new object, but it still exists as a level. Let's get rid of it.

##Remove unused factor levels

Using `fct_drop` let's remove the unused factor level:

```{r}
dropO <- noO %>% #make a new variable
  mutate(continent = fct_drop(continent)) #drop unused factors of continent

unique(dropO$continent) #check levels of continent in the new variable
```

Now we can see that Oceania was removed as a factor, yay!

##Reorder the levels of continent

For this part I will reorder the levels of continent by minimum life expectancy using `fct_reorder`.

```{r}
gapminder %>% 
  #Make a plot, reordering continent by minimum life expectancy
  ggplot(aes(fct_reorder(continent, lifeExp, .fun = min), lifeExp, fill = continent)) +
  #Add a violin plot
  geom_violin() +
  #Add an x label
  xlab("Continent") +
  #Add a y label
  ylab("Life expectancy (years)") +
  #Remove the legend
  guides(fill = FALSE)
```

Yay! We reordered the continents by minimum life expectancy.

#Part 2: File I/O

For this section I will create a new dataset by filtering the gapminder dataset for only Oceania, reordering by population size, write the file as a .csv to my local drive, and read the .csv back into R.

```{r}
OceanaData <- gapminder %>% #make a new dataset
  filter(continent == "Oceania") %>% #filter for Oceana
  arrange(pop) #arrange by population

write_csv(OceanaData, "OceanaData.csv") #write the new data to a csv

readOceana <- read_csv("OceanaData.csv") #read the csv back in
```

Let's look at what the data looked like before we wrote it to a csv

```{r}
OceanaData %>% 
  kable()
```

Great, it filtered for Oceania and ordered by population, like I wanted.

Now let's see what the dataset looks like after we read it back in:

```{r}
readOceana %>% 
  kable()
```

Great! It looks like the filtering and arranging survived through being written and read as a csv.

#Part 3: Visualization design

In this section I will recreate a graph I made in hw01. 

My graph from hw01 looked like this:

```{r}
plot(gapminder$continent, gapminder$lifeExp)
```

Let's see how we could clean this up. 

- Use `ggplot2` instead of base R graphics
- Use violin plots over boxplots to better portray the shape of the data
- Reorder your data in a meaningful way, for this I ordered by median life expectancy
- Colour each continent differently
- Add meaningful x and y labels, and a title
- Remove unnecessary information, in this case, the legend
- Add a theme, I chose `minimal` to keep it clean, and then added a background of `ghostwhite` :ghost: spooky

```{r}
#Store graph as variable `l`
l <- (gapminder %>%
#Plot continent vs lifeExp, reorder by lifeExp
ggplot(aes(fct_reorder(continent, lifeExp, .fun = median), lifeExp, fill = continent)) +
#Add a violin plot
  geom_violin() +
  #Add an x label
  xlab("Continent") +
  #Add a y label
  ylab("Life expectancy (years)") +
  #Add a title
  ggtitle("Life expectancy by continent (1952-2007)") +
  #Remove the legend
  guides(fill = FALSE) +
  #Add a theme
  theme_minimal() +
  #Change font size of axis text
  theme(axis.text = element_text(size=10),
        #Change panel background colour
        panel.background = element_rect(fill="ghostwhite")))
#Print graph
l
```

Now we will make this graph using `plotly`

```{r}
#Make plotly graph with object `l`
gg <- ggplotly(l)
#Print plotly graph
gg
#Save plotly as html, commented out so plotly could be attached as html
htmlwidgets::saveWidget(gg, file = "gg.html")
```

